---
title: "ER01-\"project\""
author: "William Aufort"
date: "1 December 2016"
output: html_document
---

# Introduction

This R Markdown document summarizes the work I've done for the ER01, which is a research school about scientific methodology and performance evaluation in computer science.

### Markdown ?

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. You can also include several contents, such a datas or code for your prefered language! For more details on using R Markdown see [this website](http://rmarkdown.rstudio.com>).

### The subject

My project deals with convex layers of 2D sets of integer points. The origin of this problem cames from arithmetic circuits, more precisely in the [tau conjecture](https://hal.inria.fr/ensl-00850791v2/document) for Newton polygons which, if we are able to prove it, implies that the permanent polynomial is not computable by polysize arithmetic circuits.

The convex layers of a set of points are the convex polygons obtained by successively "peeling" the convex hull of the set. I'm interested in some properties of this "peeling" process, for example : how many layers does a set of $n$ points can have ? How many points can we have on some particular layer ?

In this project, we are going to investigate only the first question. One can show that, if the $n$ points are uniformly distributed in a disk, then the expected number of convex layers is $\Theta(n^{2/3})$. But with this property should be true also for integer set of points ?

# The experiment

Our goal is to gather measures of convex layers for "random" set of points. But since there is no uniform distribution on the natural number, how can we modelise this?

Instead, I choose to pick all the points uniformly in a $[0,N]\times[0,N]$ square of fixed size. The wanted number of points is handled by hand, i.e selected by the program. For sake of simplicity, I allow points to appear several times in this process. Consequently, to measure the real number of points in our set, we have to take this into account in the experiment.

As a conclusion, we can summarize the experiment as follows:

1. Pick a value for $N$ the size of the square.
2. For some values of $t$ the number of points do
3. Build $f(t)$ sets of $t$ points, and compute the number of layers of the set and the real number of points.
4. Put them all in a file.

The number of sets we build will be a function of $t$, simply because the combinatorial complexity of the possible sets increases with $t$. The function $f$ we should choose is not so obvious... For sake of simplicity (and lack of time) I choose $f(t) = t^2$.

In the `convexLayers.cpp` file, you can find the code of all the experiment, with all the geometrical part of the program, i.e. compute the layers (and some plotting functions to control what we are doing). For this geometrical part, I used the [DGtal library](http://dgtal.org/) for digital geometry.

# Results

To exploit the data coming from the experiment, I use... R! What a surprise, that's the goal of the research school.
More precisely e first load the data using:

```{r}
values <- read.csv("values.csv", header = T);
```

Let's see some of them:

```{r, echo=FALSE}
head(values)
library("ggplot2")
```

These one are not so interesting, because we only see sets with few points.
Let's see what's happening with more complex sets:

```{r}
ggplot(data=values, aes(monomials, layers)) + geom_point()
```

One shouldn't be surprised by the spots on the graphic: we are working with integers! The spots are a little bit wide horizontally, because of the story about the real number of different points when we sample. 
Concerning the number of layers, it seems to increase slowly with the number of points, but to determine if we have a polyrelation between the two quantities, we can use a log-representation of the data and build a linear regression in order to determine f such a relation exists. I use the following code:

```{r}
linear_equation <- function(df){
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2, 
         list(a = format(coef(m)[1], digits = 2), 
              b = format(coef(m)[2], digits = 2), 
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));                 
}

values$x = log(values$monomials);
values$y = log(values$layers);
p <- ggplot(data=values, aes(x,y)) + geom_point() + geom_smooth(method="lm")
p + geom_text(data=NULL, "x" = 5, "y" = 1.7, label = linear_equation(values), parse = TRUE)
```

We can notice that the slope of the regression line est 0.67, which is quite close to the $2/3$ the theory gives us for uniform points in the disk, which is against our initial intuition.

# Conclusion

As a conclusion, our experiment about the number of convex layers of sets of integer points and its interpretation using R convice us that the asymptotics for integer points sets and for general points sets (in $\Theta(n^{3/2})$ should be the same. Whereas a proof for the uniform case exists, such a proof for our case is unknown, because of the combinatorial counterpart imposed by the integer coordinates.
